maze-dataset: Maze Generation with Algorithmic Variety and Representational Flexibility

### 2025-10-12 docs/paper/figures/tex/fig2_formats.tex
JOSS paper fixes
see https://github.com/openjournals/joss-reviews/issues/8633#issuecomment-3270514168

- `numpy` -> NumPy in figure 2
- proper capitalization of "feature-rich" in figure 2
- fixed legend in figure 7
- author name consistency: "Alex F. Spies" -> "Alexander F. Spies"

\begin{figure}[H]
	\centering
	\begin{tabular}{p{1.5in} p{1.5in} p{1.5in}} 
	  \hline \\[.5em]
	  % algorithms
	  \docslink{maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
	  & \docslink{maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}
	  & \docslink{maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot()}} \\[.5em]
	  % descriptions
		Simple text format for displaying mazes, useful for debugging in a terminal environment.
		& \texttt{numpy} array of \texttt{dtype=uint8} and shape \texttt{(height, width, 3)}. The last dimension is RGB color.
		& feature-rich plotting utility with support for multiple paths, heatmaps over positions, and more. \\[1em]
	  \hline \\
	  % examples
		\multicolumn{1}{c}{\begin{minipage}[b]{1.6in}
		  \setlength{\baselineskip}{0.9em}
		  \input{figures/outputs-ascii-colored.tex} 
		\end{minipage}}
		& \multicolumn{1}{c}{
		  \includegraphics[width=0.25\textwidth]{figures/outputs-pixels.pdf}
		}
		& \multicolumn{1}{c}{
		  \includegraphics[width=0.27\textwidth, trim={0 0.8cm -.3cm, -.5cm}, clip]{figures/outputs-mazeplot.pdf}
		} \\[1em]

>>>>

\begin{figure}[H]
	\centering
	\begin{tabular}{p{1.5in} p{1.5in} p{1.5in}} 
	  \hline \\[.5em]
	  % algorithms
	  \docslink{maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
	  & \docslink{maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}
	  & \docslink{maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot()}} \\[.5em]
	  % descriptions
		Simple text format for displaying mazes, useful for debugging in a terminal environment.
		& NumPy array of \texttt{dtype=uint8} and shape \texttt{(height, width, 3)}. The last dimension is RGB color.
		& Feature-rich plotting utility with support for multiple paths, heatmaps over positions, and more. \\[1em]
	  \hline \\
	  % examples
		\multicolumn{1}{l}{
		  \hspace{-1.5em}\raisebox{0.85\height}{
			  \input{figures/outputs-ascii-colored.tex}
		  }
		}
		% 	\begin{minipage}[b]{1.6in}
		%   \setlength{\baselineskip}{0.9em}
		% 	\end{minipage}
		& \multicolumn{1}{c}{
		  \includegraphics[width=0.25\textwidth]{figures/outputs-pixels.pdf}
		}
		& \multicolumn{1}{r}{
		  \includegraphics[width=0.27\textwidth, trim={0 0.8cm -.3cm, -.5cm}, clip]{figures/outputs-mazeplot.pdf}
		} \\[1em]

### 2025-10-12 docs/paper/paper.tex
reorder sections, implementation before benchmarks
JOSS paper fixes
see https://github.com/openjournals/joss-reviews/issues/8633#issuecomment-3270514168

- `numpy` -> NumPy in figure 2
- proper capitalization of "feature-rich" in figure 2
- fixed legend in figure 7
- author name consistency: "Alex F. Spies" -> "Alexander F. Spies"

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage[rgb,svgnames]{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage[pdfa]{hyperref}
\usepackage{hyperxmp}
\hypersetup{%
    unicode=true,
    pdfapart=3,
    pdfaconformance=B,
    pdftitle={maze-dataset: Maze Generation with Algorithmic Variety and
Representational Flexibility},
    pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
Spies, Tilman Räuker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
Fung},
    pdfpublication={Journal of Open Source Software},
    pdfpublisher={Open Journals},
    pdfissn={2475-9066},
    pdfpubtype={journal},
    pdfvolumenum={},
    pdfissuenum={},
    pdfdoi={10.xxxxxx/draft},
    pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
Aaron Sandoval, Alex F. Spies, Tilman Räuker, Brandon Knutson, Cecilia
Diniz Behn, Samy Wu Fung},
    pdflicenseurl={http://creativecommons.org/licenses/by/4.0/},
    colorlinks=true,
    linkcolor=[rgb]{0.0, 0.5, 1.0},
    citecolor=Blue,
    urlcolor=[rgb]{0.0, 0.5, 1.0},
    breaklinks=true
}
% https://tex.stackexchange.com/a/535849
% Create an OutputIntent in order to correctly specify colours
\immediate\pdfobj stream attr{/N 3} file{sRGB.icc}
\pdfcatalog{%
  /OutputIntents [
    <<
      /Type /OutputIntent
      /S /GTS_PDFA1
      /DestOutputProfile \the\pdflastobj\space 0 R
      /OutputConditionIdentifier (sRGB)
      /Info (sRGB)
    >>
  ]
}
\pdfvariable omitcidset=1
\usepackage{caption}
\usepackage{orcidlink}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{xstring}

@@

\author[1%
%
\ensuremath\mathparagraph]{Michael Igorevich Ivanitskiy%
  \,\orcidlink{0000-0002-4213-4993}\,%
}
\author[4%
%
]{Aaron Sandoval%
  \,\orcidlink{0009-0002-8380-6140}\,%
}
\author[2%
%
]{Alex F. Spies%
  \,\orcidlink{0000-0002-8708-1530}\,%
}
\author[3%
%
]{Tilman Räuker%
  \,\orcidlink{0009-0009-6321-4413}\,%
}
\author[1%
%
]{Brandon Knutson%
  \,\orcidlink{0009-0004-8413-0239}\,%
}
\author[1%
%
]{Cecilia Diniz Behn%
  \,\orcidlink{0000-0002-8078-5105}\,%
}
\author[1%
%
]{Samy Wu Fung%
  \,\orcidlink{0000-0002-2926-4582}\,%
}

@@

  \begin{itemize}
  \tightlist
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
    utilizing
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
    which can exactly mimic the outputs provided in
    \texttt{easy-to-hard-data}
    (\citeproc{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta, Bansal,
    et al., 2021}) and can be configured to be similar to the outputs of
    Németh (\citeproc{ref-gh_Nemeth_2019}{2019})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
    provides a format similar to
    (\citeproc{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
    \citeproc{ref-eval-gpt-visual}{Singla, 2023})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
    provides a feature‑rich plotting utility with support for multiple
    paths, heatmaps over positions, and more. This is similar to the
    outputs of (\citeproc{ref-mazegenerator-net}{Alance AB, 2019};
    \citeproc{ref-gh_Ehsan_2022}{Ehsan, 2022};
    \citeproc{ref-mathematica-maze}{Guo et al., 2011};
    \citeproc{ref-mdl-suite}{Nag, 2020})
  \end{itemize}
\item
  The text format provided by
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
  is similar to that of (\citeproc{ref-eval-LLM-graphs}{Liu \& Wu,
  2023}) but with many more options, detailed in
  \hyperref[sec:tokenized-output-formats]{section: \textit{\nameref{sec:tokenized-output-formats}}}.
\item
  Preserving metadata about the generation algorithm with the dataset
  itself is essential for studying the effects of distributional shifts.
  Our package efficiently stores the dataset along with its metadata in
  a single human-readable file (\citeproc{ref-zanj}{M. Ivanitskiy,
  n.d.}). As far as we are aware, no existing packages do this reliably.
\item
  Storing mazes as images or adjacency matrices is not only difficult to
  work with, but also inefficient. We use a highly efficient method
  detailed in
  \hyperref[sec:implementation]{section: \textit{\nameref{sec:implementation}}}.
\item
  Our package is easily installable with source code freely available.
  It is extensively tested, type hinted, benchmarked, and documented.
  Many other maze generation packages lack this level of rigor and
  scope, and some (\citeproc{ref-ayaz2008maze}{Ayaz et al., 2008})
  appear to simply no longer be accessible.
\end{itemize}

>>>>

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage[rgb,svgnames]{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage[pdfa]{hyperref}
\usepackage{hyperxmp}
\hypersetup{%
    unicode=true,
    pdfapart=3,
    pdfaconformance=B,
    pdftitle={maze-dataset: Maze Generation with Algorithmic Variety and
Representational Flexibility},
    pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alexander
F. Spies, Tilman Räuker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
Fung},
    pdfpublication={Journal of Open Source Software},
    pdfpublisher={Open Journals},
    pdfissn={2475-9066},
    pdfpubtype={journal},
    pdfvolumenum={},
    pdfissuenum={},
    pdfdoi={10.xxxxxx/draft},
    pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
Aaron Sandoval, Alexander F. Spies, Tilman Räuker, Brandon Knutson,
Cecilia Diniz Behn, Samy Wu Fung},
    pdflicenseurl={http://creativecommons.org/licenses/by/4.0/},
    colorlinks=true,
    linkcolor=[rgb]{0.0, 0.5, 1.0},
    citecolor=Blue,
    urlcolor=[rgb]{0.0, 0.5, 1.0},
    breaklinks=true
}
% https://tex.stackexchange.com/a/535849
% Create an OutputIntent in order to correctly specify colours
\immediate\pdfobj stream attr{/N 3} file{sRGB.icc}
\pdfcatalog{%
  /OutputIntents [
    <<
      /Type /OutputIntent
      /S /GTS_PDFA1
      /DestOutputProfile \the\pdflastobj\space 0 R
      /OutputConditionIdentifier (sRGB)
      /Info (sRGB)
    >>
  ]
}
\pdfvariable omitcidset=1
\usepackage{caption}
\usepackage{orcidlink}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{xstring}

@@

\author[1%
%
\ensuremath\mathparagraph]{Michael Igorevich Ivanitskiy%
  \,\orcidlink{0000-0002-4213-4993}\,%
}
\author[4%
%
]{Aaron Sandoval%
  \,\orcidlink{0009-0002-8380-6140}\,%
}
\author[2%
%
]{Alexander F. Spies%
  \,\orcidlink{0000-0002-8708-1530}\,%
}
\author[3%
%
]{Tilman Räuker%
  \,\orcidlink{0009-0009-6321-4413}\,%
}
\author[1%
%
]{Brandon Knutson%
  \,\orcidlink{0009-0004-8413-0239}\,%
}
\author[1%
%
]{Cecilia Diniz Behn%
  \,\orcidlink{0000-0002-8078-5105}\,%
}
\author[1%
%
]{Samy Wu Fung%
  \,\orcidlink{0000-0002-2926-4582}\,%
}

@@

  \begin{itemize}
  \tightlist
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
    utilizing
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
    which can exactly mimic the outputs provided in
    \texttt{easy-to-hard-data}
    (\citeproc{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta, Bansal,
    et al., 2021}) and can be configured to be similar to the outputs of
    Németh (\citeproc{ref-gh_Nemeth_2019}{2019})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
    provides a format similar to
    (\citeproc{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
    \citeproc{ref-eval-gpt-visual}{Singla, 2023})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
    provides a feature-rich plotting utility with support for multiple
    paths, heatmaps over positions, and more. This is similar to the
    outputs of (\citeproc{ref-mazegenerator-net}{Alance AB, 2019};
    \citeproc{ref-gh_Ehsan_2022}{Ehsan, 2022};
    \citeproc{ref-mathematica-maze}{Guo et al., 2011};
    \citeproc{ref-mdl-suite}{Nag, 2020})
  \end{itemize}
\item
  The text format provided by
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
  is similar to that of (\citeproc{ref-eval-LLM-graphs}{Liu \& Wu,
  2023}) but with many more options, detailed in
  \hyperref[sec:tokenized-output-formats]{section: \textit{\nameref{sec:tokenized-output-formats}}}.
\item
  Preserving metadata about the generation algorithm with the dataset
  itself is essential for studying the effects of distributional shifts.
  Our package efficiently stores the dataset along with its metadata in
  a single human-readable file (\citeproc{ref-zanj}{M. Ivanitskiy,
  n.d.}). As far as we are aware, no existing packages do this reliably.
\item
  Storing mazes as images or adjacency matrices is not only difficult to
  work with, but also inefficient. We use a highly efficient method
  detailed in
  \hyperref[sec:implementation]{section: \textit{\nameref{sec:implementation}}}.
\item
  Our package is easily installable with source code freely available.
  It is extensively tested, type hinted, benchmarked, and documented.
  Many other maze generation packages lack this level of rigor and
  scope, and some (\citeproc{ref-ayaz2008maze}{Ayaz et al., 2008})
  appear to simply no longer be accessible.
\end{itemize}
